---
title: "Phyllotaxis"
output: html_document
fig_width: 4 
fig_height: 4
---
```{r}
knitr::opts_chunk$set(fig.width=4, fig.height=4) 
```

[online tutorial][0] to take it step by step in R

# Loading ggplot2 
```{r}
# adjusting plot sizing
options(repr.plot.width = 5, repr.plot.height = 5)
library(ggplot2)
```

[more to know about fractal shapes][1]
[some freactal shapes][2]  

***

# Drawing circles
```{r}
# use 50 points to draw a circle of radius 1

t <- seq(0, 2*pi, length.out = 50) # 50 data points from 0 up to 2*pi (angles)
x <- sin(t)
y <- cos(t)
df <- data.frame(t, x, y) # data set of angles, sin and cos

# Make a scatter plot of points in a circle
p <- ggplot(df, aes(x, y))

```

## circle scatter plot
Here we take the points in df mapped using aes(x,y) & draw geometry using the geom_ functions  

* we use geom_point() here to draw just 50 points on the circumference  
* full circle can be obtained using geom_path()
```{r}
p + geom_point()
```

## Spirals, Golden angle, and harmonious shapes  

### simple spiral shape
To create a flower shpae we need to create a [spiral][4]
A spiral is a curve which starts from the origin and moves away from this point as it revolves around it
That's easily obtained by multiplying the x & y values in a factor. 
This factor can be t itself like:
```{r}
x1 <- t * x
y1 <- t * y
df1 <- data.frame(t, x1, y1)
p1 <- ggplot(df1, aes(x1, y1))
p1 + geom_point()
```

### More harmonuos spiral shape
for a more harmonuos spiral shape we use the [Golden Angle][5] concept which is also related to the Golden Ratio 

Golden Angle = π(3 − √5)
```{r}
# Defining the number of points
points <- 500

# Defining the Golden Angle
angle <- pi * (3 - sqrt(5))

# remake the scatter plot again
# t <- seq(1*angle, 500*angle, length.out = 500) # t <- (1:points) * angle
t <- (1:points) * angle
x <- sin(t)
y <-cos(t)
df <- data.frame(t, x, y)
p <- ggplot(df, aes(x*t, y*t))
p + geom_point()
```

### remove the unwanted items in the graph
we remove the grey background, the ticks and axes titles, the grid .. etc.
```{r}
df <- data.frame(t, x, y)

# Make a scatter plot of points in a spiral
p <- ggplot(df, aes(x*t, y*t))

p + geom_point() + theme(panel.background = element_rect(fill = "white", colour = "white"), title =  element_blank(), axis.ticks =  element_blank(), axis.text =  element_blank(), panel.grid =element_blank() )

```

### play with apperance  

here we make some changes to points size, color, and transparency (alpha)

```{r}


# Make a scatter plot of points in a spiral
p <- ggplot(df, aes(x*t, y*t))

p + geom_point(alpha = 0.5, size = 8, color = "darkgreen") + theme(panel.background = element_rect(fill = "white", colour = "white"), title =  element_blank(), axis.ticks =  element_blank(), axis.text =  element_blank(), panel.grid =element_blank())
```

### the dandelion

```{r}
p <- ggplot(df, aes(x*t, y*t))
p + geom_point(mapping = aes(size = t), alpha = 0.5, shape=8, color = "black") +
theme(panel.background = element_rect(fill = "white", colour = "white"),
      title =  element_blank(), axis.ticks =  element_blank(),
      axis.text =  element_blank(), panel.grid =element_blank(),
      legend.position = "none")
```

### sun flower

```{r}
p <- ggplot(df, aes(x*t, y*t))
p + geom_point(mapping = aes(size = t), alpha = 0.5, shape=17, color = "yellow") +
theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
      title =  element_blank(), axis.ticks =  element_blank(),
      axis.text =  element_blank(), panel.grid =element_blank(),
      legend.position = "none")
```

### Changing the angle effect

we change the golden angle, and we increase points number
we set angle to 2 with 1000 points
```{r}

    angle <- 2
    points <- 1000
    
    t <- (1:points)*angle
    x <- sin(t)
    y <- cos(t)
    
    df <- data.frame(t, x, y)
  
    p <- ggplot(df, aes(x*t, y*t))
    p + geom_point(mapping = aes(size = t), alpha = 0.5, shape=17, color = "yellow") +
    theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
          title =  element_blank(), axis.ticks =  element_blank(),
          axis.text =  element_blank(), panel.grid =element_blank(),
          legend.position = "none")
    
ggsave("2.png", width = 4, height = 4)
```

## Beautiful example

```{r}
angle <- 1.77
points <- 2000

t <- (1:points)*angle
x <- sin(t)
y <- cos(t)

df <- data.frame(t, x, y)

p <- ggplot(df, aes(x*t, y*t))
p + geom_point(mapping = aes(size = t), alpha = 0.1, shape=16, color = "pink") +
theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
      title =  element_blank(), axis.ticks =  element_blank(),
      axis.text =  element_blank(), panel.grid =element_blank(),
      legend.position = "none")

ggsave("3.png", width = 4, height = 4)
```

## another example

```{r}
angle <- 13 * pi / 180
points <- 1300

t <- (1:points)*angle
x <- sin(t)
y <- cos(t)

df <- data.frame(t, x, y)

p <- ggplot(df, aes(x*t, y*t))
p + geom_point(size = 80, alpha = 0.1, shape=1, color = "magenta4") +
theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
      title =  element_blank(), axis.ticks =  element_blank(),
      axis.text =  element_blank(), panel.grid =element_blank(),
      legend.position = "none")
ggsave("4.png", width = 3, height = 3)
```

```{r}
angle <- 19 * pi / 180
points <- 1500

t <- (1:points)*angle
x <- sin(t)
y <- cos(t)

df <- data.frame(t, x, y)

p <- ggplot(df, aes(x*t, y*t))
p + geom_point(size = 80, alpha = 0.1, shape=1, color = "Turquoise") +
theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
      title =  element_blank(), axis.ticks =  element_blank(),
      axis.text =  element_blank(), panel.grid =element_blank(),
      legend.position = "none")
ggsave("1.png", width = 4, height = 4)
```
 
### playing with a function

```{r}
play_phyllotaxis <-function(factor, points = 2000, color){
  angle <-  factor*pi / 180
  points <- points

  t <- (1:points)*angle
  x <- sin(t)
  y <- cos(t)
  
  df <- data.frame(t, x, y)
  
  p <- ggplot(df, aes(x*t, y*t))
  p + geom_point(size = 80, alpha = 0.1, shape=1, color = color) +
  theme(panel.background = element_rect(fill = "white", colour = "darkmagenta"),
        title =  element_blank(), axis.ticks =  element_blank(),
        axis.text =  element_blank(), panel.grid =element_blank(),
        legend.position = "none")
  
}
```

```{r}

play_phyllotaxis(15, 2000, "red")
ggsave("5.png", width = 4, height = 4)
```

```{r}
play_phyllotaxis(0.75, 2000, "green")
ggsave("6.png", width = 4, height = 4)
```

```{r}
play_phyllotaxis(72, 2000, "blue")
ggsave("6.png", width = 4, height = 4)
```



[0]: https://projects.datacamp.com/projects/62
[1]: https://en.wikipedia.org/wiki/Fractal
[2]: https://wallhere.com/en/wallpaper/778140
[4]: https://en.wikipedia.org/wiki/Spiral
[5]: https://en.wikipedia.org/wiki/Golden_angle